### 2024-01-11 abstract.qmd
add frontmatter

---
title: A mixture of hidden Markov models to predict the lymphatic spread on head and neck cancer
author:
  - name: Roman Ludwig
    orcid: 0000-0001-9434-328X
    email: roman.ludwig@usz.ch
    affiliations:
      - ref: 1
      - ref: 2
  - name: Julian Brönnimann
    affiliations:
      - ref: 1
      - ref: 2
  - name: Jan Unkelbach
    orcid: 0000-0002-4275-990X
    affiliations:
      - ref: 1
      - ref: 2
affiliation:
  - id: 1
    name: University of Zurich
    departement: Departement of Physics
    city: Zurich
    country: Switzerland
  - id: 2
    name: University Hospital Zurich
    departement: Departement of Radiation Oncology
    city: Zurich
    country: Switzerland
bibliography: references.bib
abstract: |
  We previously developed a mechanistic hidden Markov model (HMM) to predict the
  lymphatic tumor progression in oropharyngeal squamous cell carcinomas (OPSCCs). To
  extend the model to other tumor subsites, defined by ICD-10 codes, in the head and
  neck, we employ a mixture of these HMMs and learn the cluster assignments and model
  parameters in an iterative, EM-like algorithm from multi-centric data. The mixture
  model manages to cluster anatomically close subsites and correctly infers the
  clusters' parameters. Using this mixture model allows the prediction of individual
  risks of occult nodal disease, given a diagnosis that includes tumor subsite.
---

# Introduction

# Materials and Methods

# Results

# Discussion