### 2024-01-18 content/manuscript.qmd
add info about data & edits (see below)
I have removed the formalism related to sensitivity and specificity.
Also, the mixture model is now not motivated as a lift of the old
model's limitations anymore.

---
title: A mixture of hidden Markov models to predict the lymphatic spread in head and neck cancer
author:
  - name: Roman Ludwig
    orcid: 0000-0001-9434-328X
    email: roman.ludwig@usz.ch
    affiliations:
      - ref: 1
      - ref: 2
  - name: Julian Brönnimann
    affiliations:
      - ref: 1
      - ref: 2
  - name: Yoel Perez Haas
    affiliations:
      - ref: 1
      - ref: 2
  - name: Esmée Lauren Looman
    affiliations:
      - ref: 1
      - ref: 2
  - name: Sergi Benavente
    orcid: 0000-0002-2700-3522
    affiliations:
      - ref: 11
  - name: Adrian Schubert
    affiliations:
      - ref: 3
      - ref: 4
      - ref: 7
  - name: Dorothea Barbatei
    affiliations:
      - ref: 8
  - name: Laurence Bauwens
    affiliations:
      - ref: 8
  - name: Jean-Marc Hoffmann
    orcid: 0000-0003-1193-0863
    affiliations:
      - ref: 2
  - name: Sandrine Werlen
    affiliations:
      - ref: 4
      - ref: 5
  - name: Olgun Elicin
    orcid: 0000-0002-6996-0646
    affiliations:
      - ref: 3
  - name: Matthias Dettmer
    orcid: 0000-0003-0948-1392
    affiliations:
      - ref: 6
      - ref: 10
  - name: Philippe Zrounba
    orcid: 0000-0001-6223-6113
    affiliations:
      - ref: 9
  - name: Bertrand Poumayou
    orcid: 0000-0002-8060-7827
    affiliations:
      - ref: 2
  - name: Panagiotis Balermpas
    orcid: 0000-0001-5261-6446
    affiliations:
      - ref: 2
  - name: Vincent Grégoire
    affiliations:
      - ref: 8
  - name: Roland Giger
    orcid: 0000-0002-5574-3210
    affiliations:
      - ref: 4
      - ref: 5
  - name: Jan Unkelbach
    orcid: 0000-0002-4275-990X
    affiliations:
      - ref: 1
      - ref: 2
affiliations:
  - id: 1
    name: University of Zurich
    department: Department of Physics
    city: Zurich
    country: Switzerland
  - id: 2
    name: University Hospital Zurich
    department: Department of Radiation Oncology
    city: Zurich
    country: Switzerland
  - id: 3
    name: Bern University Hospital
    department: Department of Radiation Oncology
    city: Bern
    country: Switzerland
  - id: 4
    name: Bern University Hospital
    department: Department of ENT, Head & Neck Surgery
    city: Bern
    country: Switzerland
  - id: 5
    name: Bern University Hospital
    department: Head and Neck Anticancer Center
    city: Bern
    country: Switzerland
  - id: 6
    name: Bern University Hospital
    department: Institute of Tissue Medicine and Pathology
    city: Bern
    country: Switzerland
  - id: 7
    name: Réseau Hospitalier Neuchâtelois
    department: Department of ENT, Head & Neck Surgery
    city: Neuchâtelois
    country: Switzerland
  - id: 8
    name: Centre Léon Bérard
    department: Department of Radiation Oncology
    city: Lyon
    country: France
  - id: 9
    name: Centre Léon Bérard
    department: Department of Head and Neck Surgery
    city: Lyon
    country: France
  - id: 10
    name: Klinikum Stuttgart
    department: Institute of Pathology
    city: Stuttgart
    country: Germany
  - id: 11
    name: Hospital Vall d'Hebron
    department: Departement of Radiation Oncology
    city: Barcelona
    country: Spain
bibliography: ../references.bib
abstract: |
  We previously developed a mechanistic hidden Markov model (HMM) to predict the
  lymphatic tumor progression in oropharyngeal squamous cell carcinomas (OPSCCs). To
  extend the model to other tumor subsites, defined by ICD-10 codes, in the head and
  neck, we employ a mixture of these HMMs and learn the cluster assignments and model
  parameters in an iterative, EM-like algorithm from multi-centric data. The mixture
  model manages to cluster anatomically close subsites and correctly infers the
  clusters' parameters. Using this mixture model allows the prediction of individual
  risks of occult nodal disease, given a diagnosis that includes tumor subsite.
format:
  iccr-pdf:
    keep-tex: true
    include-in-header:
      - text: "\\usepackage{newtxtext,newtxmath}"
  html:
    code-line-numbers: true
    code-tools: true
crossref:
  fig-prefix: "figure"
  tbl-prefix: "table"
  eq-prefix: "equation"
  sec-prefix: "section"
---

# Introduction

Head and neck squamous cell carcinomas (HNSCC) frequently spread through the lymphatic system [@lindberg_distribution_1972;@woolgar_histological_1999]. Current diagnostic imaging modalities are unable to detect microscopic nodal metastases, which requires pathological examination of extracted tissue [@snyder_petct_2021;@strohl_petct_2021]. Since the recurrence of nodal disease is detrimental to a patient's outcome [@ho_decision_2014], large volumes in the head and neck region are often irradiated electively to minimize the risk of missing occult disease. Decision guidelines about which nodal regions -- i.e., anatomically defined lymph node levels (LNLs) -- to irradiate [@biau_selection_2019] are currently not based on a patient's individual risk, but only on the overall prevalence of nodal disease as reported in the literature [@lindberg_distribution_1972;@woolgar_histological_1999].

To personalize this prediction of the risk for occult diease, given a patient's individual diagnosis, we published

1. large, multi-centric data that reports per patient which LNLs where clinically and/or pathologically involved [@ludwig_dataset_2022;@ludwig_multi-centric_2023].

And, building on this work,

2. an interpretable hidden Markov model (HMM), trained with this data, to predict the risk for occult nodal disease [@ludwig_hidden_2021], given an individual patient's diagnosis.

Such a personalized risk prediction may allow clinitians to safely reduce the elective clinical target volume (CTV-N) and thus reduce side-effects that degrade the patient's quality of life [@batth_practical_2014].

HNSCC patients with primary tumors at different subsites, e.g. in the oral cavity and in the oropharynx, also show different patterns of lymphatic spread [@lindberg_distribution_1972;@woolgar_histological_1999]. Our model does so far not have the capability to naturally describe different tumor subsites. To that end, we present an approach using mixtures of these HMMs in this work. This makes intuitive sense, because if a tumor lies anatomically between two anatomically close subsites with slightly different spread patterns, we may be able to describe its lymphatic progression as a mixture of these two spread patterns.

# Materials and Methods {#sec-materials-and-methods}

To compute the personalized risk of occult disease $\mathbf{X}$, given a diagnosis $\mathbf{Y}$, we can begin by stating Bayes' law:

$$
P \left( \mathbf{X} \mid \mathbf{Y} \right) = \frac{P \left( \mathbf{Y} \mid \mathbf{X} \right) P \left( \mathbf{X} \right)}{\sum_{\mathbf{X}^\star} P \left( \mathbf{Y} \mid \mathbf{X}^\star \right) P \left( \mathbf{X}^\star \right)}
$$ {#eq-bayes-law}

In the above equation, the term $P \left( \mathbf{Y} \mid \mathbf{X} \right)$ is given by the sensitivity and specificity of the diagnosis. The crucial term that our model attempts to compute, is the prior probability of involvement $P \left( \mathbf{X} \right)$ (or rather of the hidden state of involvement, see the [next section](#sec-hmm)).

## Hidden Markov Model for Lymphatic Progression {#sec-hmm}

Each LNL $v \in \{ 1, 2, \ldots, V\}$ considered in our model is represented by a binary random variable (RV) $X_v[t]$ taking on the true state of that level at the abstract time-step $t$ (0 for "healthy" and 1 for "involved"). Collected in a random vector $\mathbf{X}[t] = \left( X_1[t], X_2[t], \ldots, X_V[t] \right)$ they form the patient's state w.r.t. their lymphatic involvement at time $t$.

We model the process of tumor progression via lymphatic drainage by connecting the RVs in a graph, as shown in @fig-graph-with-obs. The arcs in [this graph](#fig-graph-with-obs) represent conditional probabilities. The orange arcs correspond to observing a diagnosis $Y_v$, given the true state $X_v$. For the sake of brevity, we will not go into the details of describing how to infer the true -- but technically hidden -- state of LNL involvement from diagnoses with lower-than-one sensitivity and specificity. This description can be found in @zora231470. Throughout this work, we instead combine diagnostic and pathologic involvement information from the data into a "maximum likelihood" diagnosis and assume its sensitivity and specificity to be one, meaning the normally hidden state $X_v$ becomes the observed state. This simplification is reasonable for pathologic involvement and because we are at this stage more interested in testing whether our model is able to describe a realistic probability distribution over lymphatic involvement.

The red arcs in the graph of @fig-graph-with-obs depict the probability that the primary tumor spreads within one abstract time-step. While the blue arcs symbolize the spread from an upstream LNL -- given it is already metastatic -- to a downstream level. For example, the edge from $X_2$ to $X_3$ encodes the probability $P\left( X_3[t+1] \mid X_2[t] \right)$, which is parametrized with $b_3$, and $t_{2 \rightarrow 3}$ and tabulated in @tbl-spread-2-to-3. There is an additional restriction on any LNL's state $X_v[t+1]=0$ to be healthy: It requires that the level was also healthy in the time-step before, meaning $X_v[t]=0$. This is because we assume no spontaneous self-healing of metastatic levels.

+--------------+------------+-----------------------------------------------------+
|              | $X_2[t]=0$ | $X_2[t]=1$                                          |
+==============+:==========:+:===================================================:+
| $X_3[t+1]=0$ | $1-b_3$    | $(1-b_3)(1-t_{2 \rightarrow 3})$                    |
+--------------+------------+-----------------------------------------------------+
| $X_3[t+1]=1$ | $b_3$      | $1-b_3-t_{2 \rightarrow 3}+b_3 t_{2 \rightarrow 3}$ |
+--------------+------------+-----------------------------------------------------+

: Conditional probability $P\left( X_3[t+1] \mid X_2[t] \right)$ for a spread from LNL II to III during the transition from $t$ to $t+1$. This corresponds to one of the blue arcs in @fig-graph-with-obs. Note that the values in the row with $X_3[t+1]=0$ is all zeros, and the row with $X_3[t+1]=1$ all ones if $X_3[t]=1$. {#tbl-spread-2-to-3 tbl-colwidths="[25,18,40]"}

With the introduced conditional probabilities, we can now compute the joint probability of any complete state $\mathbf{X}[t]=\boldsymbol{\xi}_i$ transitioning to any other possible state $\mathbf{X}[t+1]=\boldsymbol{\xi}_j$. Here, when we use $\boldsymbol{\xi}$ instead of $\mathbf{x}$ for the values the random vector $\mathbf{X}$ can take on, the $i$ and $j$ enumerate all $2^V$ combinations of the $V$ binary RVs. In the graph shown in @fig-graph-with-obs, this amounts to $2^V=8$ distinct $\boldsymbol{\xi}_i$. Because these terms are essentially products of terms like those in @tbl-spread-2-to-3. We can then collect these terms in a _transition matrix_ $\mathbf{A}$:

$$
\mathbf{A} = \begin{pmatrix} A_{ij} \end{pmatrix} = P \left( \mathbf{X}[t+1] = \boldsymbol{\xi}_j \mid \mathbf{X}[t] = \boldsymbol{\xi}_i \right)
$$ {#eq-transition-matrix}

Note that this matrix still depends on the $b_v$ and $t_{r \rightarrow v}$ parameters, although we have dropped the explicit dependcy to keep the equations brief. Now, assuming that every patient started their disease with all LNLs being healthy, we can define the _starting distribution_ $\boldsymbol{\pi}$:

$$
\boldsymbol{\pi} = \begin{pmatrix} \pi_i \end{pmatrix} = P \left( \mathbf{X}[0] = \boldsymbol{\xi}_i \right)
$$ {#eq-starting-distribution}

And set every entry of this starting distribution to zero, except the first one, which we set to one. This means at $t=0$ there is a probability of one to be in the completly healthy state $\boldsymbol{\xi}_0 = \begin{pmatrix} 0, 0, \ldots, 0 \end{pmatrix}$.

Using the quantities introduced so far, the probability distribution vector with elements $P \left( \mathbf{X}[t]=\boldsymbol{\xi}_i \right)$ after $t$ time-steps can now be conveniently expressed as a matrix product:

$$
P \left( \mathbf{X}[t]=\boldsymbol{\xi}_i \right) = \left( \boldsymbol{\pi} \cdot \mathbf{A}^t \right)_i
$$ {#eq-evolution}

This evolution implicitly marginalizes over all possible paths to arrive at state $\boldsymbol{\xi}_i$ after $t$ time-steps. Additionally, we also need to marginalize over the time of diagnosis -- which is unknown -- using a time-prior $P_T(t)$. Fortunately, the exact length and shape of this distribution on its own has little impact. But because we assume that early and advanced T-category patients are fundamentally the same, just on average diagnosed at different times $t$, we use the time-prior only to separate the respective patient's evolutions:

$$
P \left( \mathbf{X}=\boldsymbol{\xi}_i \mid T, \boldsymbol{\theta} \right) = \sum_{t=0}^{t_\text{max}} P_T(t) \left( \boldsymbol{\pi} \cdot \mathbf{A}^t \right)_i
$$ {#eq-marginalized-evolution}

where $T \in \{ \text{early}, \text{advanced} \}$ denotes the T-category. We fix $P_\text{early}(t)$ to a binomial distribution with parameters $n=t_\text{max}=$ {{< var t_max >}} and $p_\text{early}=$ {{< var time-prior.early.p >}}, while the advanced T-category's time-prior is also a binomial distribution where the $p_\text{advanced}$ parameter is learned.

Note that @eq-marginalized-evolution still depends on the parametrization of the transition matrix. We collect these parameters in a vector $\boldsymbol{\theta}=\{ b_v, t_{r \rightarrow v}, p_\text{advanced} \}$.

To train our model, we need to compute the likelihood of a dataset $\mathbf{D}=\left( \mathbf{x}_1, \mathbf{x}_2, \ldots, \mathbf{x}_N \right)$ of $N$ patients, given the model's parameters $\boldsymbol{\theta}$. This is given by a product of the terms in @eq-marginalized-evolution:

$$
P \left(\mathbf{D} \mid \boldsymbol{\theta} \right) = \prod_{i=1}^N P \left( \mathbf{X} = \mathbf{x}_i \mid T_i, \boldsymbol{\theta} \right)
$$ {#eq-likelihood}

Typically, one would compute the logarithm of this quantity for computational reasons. There is a wide array of available methods to obtain maximum likelihood estimates from this function or sample from the posterior over $\boldsymbol{\theta}$. In this work, we use Markov chain Monte Carlo sampling (MCMC) via the `emcee` Python package [@foreman-mackey_emcee_2013].

![Parametrized graph representation of the lymphatic network considering four LNLs. Blue, round nodes represent the hidden RVs, while orange square nodes show the observed RVs. Arcs represent a conditional probability parametrized with the quantity noted next to it.](../static/graph_with_obs.svg){#fig-graph-with-obs width="25%" fig-pos="th"}

## Mixture of HMMs {#sec-mixture-of-hmms}

The just introduced model is capable of learning one set of or distribution over parameters $\boldsymbol{\theta}$ from a cohort of patients with a primary tumor in a given subsite. If we tried to train it with a cohort consisting of patients with tumors in two very different subsites, the model would likely learn parameters that represent a compromise between the two subcohort's true parameters. This compromise might describe neither of the subcohorts' lymphatic spread patterns sufficiently well.

In such cases, mixture models are often considered. They assume the data to come from a finite mixture distribution, which -- in our particular case -- can be written as follows:

$$
P \left( \mathbf{D} \mid \boldsymbol{\Psi} \right) = \sum_{j=0}^g c_j P \left( \mathbf{D} \mid \boldsymbol{\theta}_j \right)
$$ {#eq-mixture-distribution}

Here, the $\mathbf{c} \in [0,1]^g$ is the vector of mixing proportions with $\sum_{j=0}^g c_j = 1$, while the $\boldsymbol{\Psi} = \left( \boldsymbol{\theta}_1, \ldots, \boldsymbol{\theta}_g\right)$ is the vector of all $g$ models' parameters. Note that we will implement our model such that some of the parameters in each $\boldsymbol{\theta}_j$ are shared across the $g$ components -- namely the $t_{r \rightarrow v}$ corresponding to the blue arcs in @fig-graph-with-obs.

Let now $\mathfrak{D} = \left( \mathbf{D}_1, \ldots, \mathbf{D}_s \right)$ be a dataset consisting of $s$ subcohorts of patients. Within a subcohort $i$ we find $N_i$ patients with tumors in the same subsite. We can then introduce a _latent variable_ $\mathbf{Z}$ with a one-hot-encoding: Basically, it can take on values $\mathbf{z}_i \in \{ 0,1 \}^g$ with $z_{ij}=1$ if subcohort $i$ belongs to component $j$ and $z_{ij}=0$ else.

The latent variables are helpful in resolving the invariance of the likelihood w.r.t. permutations of the component labels, which may introduce problems, e.g. for common MCMC sampling methods. The $\mathbf{Z}$ allows us to derive two sets of interdependent equations that we may solve in an iterative fashion (see e.g. @bishop_pattern_2006 for a detailed derivation) that is commonly referred to as _expectation-maximization (EM)_ algorithm:

The first set are the probabilities of subcohort $i$ to belong to component $j$, given a set of parameters $\boldsymbol{\Psi}^\star$. These are often called the _responsibilities_:
$$
\gamma (z_{ij}) = P \left( z_{ij}=1 \mid \mathbf{D}_i , \boldsymbol{\Psi}^\star, \mathbf{c} \right) = \frac{ c_j P \left( \mathbf{D}_i \mid \boldsymbol{\theta}_j^\star \right) }{\sum_{k=0}^g c_k P \left( \mathbf{D}_i \mid \boldsymbol{\theta}_k^\star \right)}
$$

From this, we can compute new mixing proportions $c_j^\star=\sum_{i=1}^s \gamma (z_{ij}) / s$ and then infer new parameters $\boldsymbol{\Psi}^\star$ -- e.g. via MCMC sampling -- from the resulting likelihood, which is the second set:

$$
P \left( \mathbf{D} \mid \boldsymbol{\Psi}, \mathbf{c}^\star \right) = \sum_{j=0}^g c_j^\star P \left( \mathbf{D} \mid \boldsymbol{\theta}_j \right)
$$

## Implementation {#sec-implementation}

We did this in this and that fashion...

## Multi-Centric Data

For the analyses in this work, we used five datasets from four different institutions:

1. 287 oropharyngeal patients from the University of Zurich in Swizerland
2. 263 oropharyngeal patients from the Centre Léon Bérard in Fance
3. 289 oropharyngeal and oral cavity patients from the Inselspital Bern in Swizerland
4. 239 oropharyngeal and oral cavity patients from the Centre Léon Bérard in Fance
5. 162 oropharyngeal patients from the Hospital Vall d'Hebron in Spain

The data comes in the form of CSV tables and are -- except for the last and most recent addition -- publicly available [@ludwig_detailed_2022;@ludwig_multi-centric_2023] and may be interactively explored in our **Ly**mphatic **Pro**gression e**X**plorer [LyProX](https://lyprox.org). Each row of these tables corresponds to one patient and details in which LNL metastatic involvement was found or not, according to different diagnostic and pathologic modalities.

# Results {#sec-results}

Works super well, of course!

# Discussion {#sec-discussion}

All that's left is for other people to bring this into clinical practice.