### 2024-01-25 content/manuscript.qmd
work in Jan's edits

---
title: A mixture of hidden Markov models to predict the lymphatic spread in head and neck cancer
author:
  - name: Roman Ludwig
    orcid: 0000-0001-9434-328X
    email: roman.ludwig@usz.ch
    affiliations:
      - ref: 1
      - ref: 2
  - name: Julian Brönnimann
    affiliations:
      - ref: 1
      - ref: 2
  - name: Yoel Perez Haas
    affiliations:
      - ref: 1
      - ref: 2
  - name: Esmée Lauren Looman
    affiliations:
      - ref: 1
      - ref: 2
  - name: Sergi Benavente
    orcid: 0000-0002-2700-3522
    affiliations:
      - ref: 11
  - name: Adrian Schubert
    affiliations:
      - ref: 3
      - ref: 4
      - ref: 7
  - name: Dorothea Barbatei
    affiliations:
      - ref: 8
  - name: Laurence Bauwens
    affiliations:
      - ref: 8
  - name: Jean-Marc Hoffmann
    orcid: 0000-0003-1193-0863
    affiliations:
      - ref: 2
  - name: Sandrine Werlen
    affiliations:
      - ref: 4
      - ref: 5
  - name: Olgun Elicin
    orcid: 0000-0002-6996-0646
    affiliations:
      - ref: 3
  - name: Matthias Dettmer
    orcid: 0000-0003-0948-1392
    affiliations:
      - ref: 6
      - ref: 10
  - name: Philippe Zrounba
    orcid: 0000-0001-6223-6113
    affiliations:
      - ref: 9
  - name: Bertrand Poumayou
    orcid: 0000-0002-8060-7827
    affiliations:
      - ref: 2
  - name: Panagiotis Balermpas
    orcid: 0000-0001-5261-6446
    affiliations:
      - ref: 2
  - name: Vincent Grégoire
    affiliations:
      - ref: 8
  - name: Roland Giger
    orcid: 0000-0002-5574-3210
    affiliations:
      - ref: 4
      - ref: 5
  - name: Jan Unkelbach
    orcid: 0000-0002-4275-990X
    affiliations:
      - ref: 1
      - ref: 2
affiliations:
  - id: 1
    name: University of Zurich
    department: Department of Physics
    city: Zurich
    country: Switzerland
  - id: 2
    name: University Hospital Zurich
    department: Department of Radiation Oncology
    city: Zurich
    country: Switzerland
  - id: 3
    name: Bern University Hospital
    department: Department of Radiation Oncology
    city: Bern
    country: Switzerland
  - id: 4
    name: Bern University Hospital
    department: Department of ENT, Head & Neck Surgery
    city: Bern
    country: Switzerland
  - id: 5
    name: Bern University Hospital
    department: Head and Neck Anticancer Center
    city: Bern
    country: Switzerland
  - id: 6
    name: Bern University Hospital
    department: Institute of Tissue Medicine and Pathology
    city: Bern
    country: Switzerland
  - id: 7
    name: Réseau Hospitalier Neuchâtelois
    department: Department of ENT, Head & Neck Surgery
    city: Neuchâtelois
    country: Switzerland
  - id: 8
    name: Centre Léon Bérard
    department: Department of Radiation Oncology
    city: Lyon
    country: France
  - id: 9
    name: Centre Léon Bérard
    department: Department of Head and Neck Surgery
    city: Lyon
    country: France
  - id: 10
    name: Klinikum Stuttgart
    department: Institute of Pathology
    city: Stuttgart
    country: Germany
  - id: 11
    name: Hospital Vall d'Hebron
    department: Departement of Radiation Oncology
    city: Barcelona
    country: Spain
bibliography: ../references.bib
abstract: |
  We previously developed a mechanistic hidden Markov model (HMM) to predict the
  lymphatic tumor progression in oropharyngeal squamous cell carcinomas (OPSCCs). To
  extend the model to other tumor subsites, defined by ICD-10 codes, in the head and
  neck, we employ a mixture of these HMMs and learn the cluster assignments and model
  parameters in an iterative, EM-like algorithm from multi-centric data. The mixture
  model manages to cluster anatomically close subsites and correctly infers the
  clusters' parameters. Using this mixture model allows the prediction of individual
  risks of occult nodal disease, given a diagnosis that includes tumor subsite.
format:
  iccr-pdf:
    keep-tex: true
    include-in-header:
      - text: "\\usepackage{newtxtext,newtxmath}"
  html:
    code-line-numbers: true
    code-tools: true
crossref:
  fig-prefix: "figure"
  tbl-prefix: "table"
  eq-prefix: "equation"
  sec-prefix: "section"
---

# Introduction

Head and neck squamous cell carcinomas (HNSCC) frequently spread through the lymphatic system [@lindberg_distribution_1972;@woolgar_histological_1999]. Current diagnostic imaging modalities are unable to detect microscopic lymph node metastases [@snyder_petct_2021;@strohl_petct_2021]. To avoid nodal recurrences [@ho_decision_2014], large volumes in the neck are irradiated electively, which are at risk of harbouring occult disease. Guidelines about which lymph node levels (LNLs) to irradiate [@biau_selection_2019] are currently not based on a patient's individual risk, but only on the overall prevalence of nodal disease as reported in the literature [@lindberg_distribution_1972;@woolgar_histological_1999].

To personalize this prediction of the risk for occult disease, given a patient's individual diagnosis, we published

1. large, multi-centric data that reports per patient which LNLs where clinically and/or pathologically involved [@ludwig_dataset_2022;@ludwig_multi-centric_2023].

And, building on this work,

2. an interpretable hidden Markov model (HMM), trained with this data, to predict the risk for occult nodal disease [@ludwig_hidden_2021], given an individual patient's diagnosis.

Such a personalized risk prediction may allow clinicians to safely reduce the elective clinical target volume (CTV-N) and thus reduce side-effects that degrade the patient's quality of life [@batth_practical_2014].

Here, we extend the previous work by incorporating the primary tumor location (specified as ICD-10 code) into the model of lymphatic tumour progression, focusing on tumours in the oropharynx and the oral cavity. HNSCC patients with primary tumors at different subsites show different patterns of lymphatic spread [@lindberg_distribution_1972;@woolgar_histological_1999]. So far, this could be handled by training different models for broader categories of tumour locations, e.g. oropharynx and oral cavity tumours. However, this approach does not describe differences in lymphatic spread between different subsites within the oropharynx and oral cavity. To address this issue, we present an approach using mixtures of HMMs. The intuition is that the lymphatic spread of a tumor that lies anatomically at the boarder of oropharynx and oral cavity (e.g. tumours in the palate) may be described by a mixture of different models.

# Materials and Methods {#sec-materials-and-methods}

Each LNL $v \in \{ 1, 2, \ldots, V\}$ considered in our model is represented by a binary random variable (RV) $X_v$ representing the true state of that level (0 for "healthy" and 1 for "involved"). A patient's state of lymph node involvement can be represented in a random vector $\mathbf{X} = \left( X_1, X_2, \ldots, X_V \right)$. When a patient is diagnosed with HNSCC, we only observe the clinical lymph node involvement based on imaging, which we denote as another binary random variable $Y_v$. To compute the personalized risk of occult disease $\mathbf{X}$, given a diagnosis $\mathbf{Y}$, we apply Bayes' law:

$$
P \left( \mathbf{X} \mid \mathbf{Y} \right) = \frac{P \left( \mathbf{Y} \mid \mathbf{X} \right) P \left( \mathbf{X} \right)}{\sum_{\mathbf{X}^\star} P \left( \mathbf{Y} \mid \mathbf{X}^\star \right) P \left( \mathbf{X}^\star \right)}
$$ {#eq-bayes-law}

In the above equation, the term $P \left( \mathbf{Y} \mid \mathbf{X} \right)$ is given by the sensitivity and specificity of the diagnostic procedure. The term $P \left( \mathbf{X} \right)$ represents the prior probability of involvement, which depends on the probability of the tumour to spread through the lymphatic system. The main task of the HMM is to model $P \left( \mathbf{X} \right)$ and the main contribution of this paper is to incorporate the primary tutor subset into the model of $P \left( \mathbf{X} \right)$.

## Hidden Markov Model for Lymphatic Progression {#sec-hmm}

A patient's state of lymph node involvement $\mathbf{X}[t]$ evolves over discrete time steps $t$. Let us enumerate all $2^V$ possible states, representing all combinations of LNLs. In this paper, we consider ipsilateral LNLs I, II, III and IV, which amounts to 16 possible states. The HMM is specified by _transition matrix_ $\mathbf{A}$:

$$
\mathbf{A} = \begin{pmatrix} A_{ij} \end{pmatrix} = P \left( \mathbf{X}[t+1] = \boldsymbol{\xi}_j \mid \mathbf{X}[t] = \boldsymbol{\xi}_i \right)
$$ {#eq-transition-matrix}

which contains the conditional probabilities that a state $\mathbf{X}[t]=\boldsymbol{\xi}_i$ transitions to $\mathbf{X}[t+1]=\boldsymbol{\xi}_j$ over one time step. The transition matrix is specified and parameterised via the graphical model shown in @fig-schematic-and-graph. The red arcs in the graph of @fig-schematic-and-graph are associated the probability that the primary tumor spreads directly to a LNL (parameters $b_v$). The blue arcs describe the spread from an upstream LNL -- given it is already metastatic -- to a downstream level (parameters $t_{v-1 \rightarrow v}$).

Now, let $\boldsymbol{\pi}$ be the _starting distribution_

$$
\boldsymbol{\pi} = \begin{pmatrix} \pi_i \end{pmatrix} = P \left( \mathbf{X}[0] = \boldsymbol{\xi}_i \right)
$$ {#eq-starting-distribution}

denoting the probability to start in state $\boldsymbol{\xi}_i$ at time step 0. Assuming that every patient started their disease with all LNLs being healthy, we set $\pi_i$ to zero for all states  except the completly healthy state $\boldsymbol{\xi} = \begin{pmatrix} 0, 0, 0, 0 \end{pmatrix}$, which has probability 1.

Using the quantities introduced so far, the probability $P \left( \mathbf{X}[t]=\boldsymbol{\xi}_i \right)$ to be in state $i$ in time step $t$ can now be conveniently expressed as a matrix product:

$$
P \left( \mathbf{X}[t]=\boldsymbol{\xi}_i \right) = \left( \boldsymbol{\pi} \cdot \mathbf{A}^t \right)_i
$$ {#eq-evolution}

This evolution implicitly marginalizes over all possible paths to arrive at state $\boldsymbol{\xi}_i$ after $t$ time-steps. Additionally, we must marginalize over the unknown time of diagnosis using a time-prior $P(t)$. This finally defines the probability distribution over all states of lymph node involvement used in @eq-bayes-law.

$$
P \left( \mathbf{X}=\boldsymbol{\xi}_i \mid \boldsymbol{\theta} \right) = \sum_{t=0}^{t_\text{max}} P_T(t) \left( \boldsymbol{\pi} \cdot \mathbf{A}^t \right)_i
$$ {#eq-marginalized-evolution}

where $\boldsymbol{\theta}=\{ b_v, t_{r \rightarrow v} \}$ denotes the set of all model parameters (7 in our case). Fortunately, the exact length and shape of this distribution on its own has little impact as previously shown. We set $t_\text{max}=$ {{< var model.max_t >}} and $P_\text{early}(t)$ to a binomial distribution with parameter {{< var model.first_binom_prob >}}. Further details on the HMM can be found in ...

![On the left: Rough anatomical sketch of the tumor subsites and correspsonding ICD-10 codes that are present in the used data. The subsite "other parts of mouth" (C06) was not drawn. On the right: Parametrized graph representation of the lymphatic network considering four LNLs. Blue nodes represent the hidden RVs, while the red one is the tumor. Arcs represent a conditional probability parametrized with the quantity noted next to it](../static/schematic_and_graph.svg){#fig-schematic-and-graph fig-pos="th"}

## Mixture of HMMs {#sec-mixture-of-hmms}

Let us now assume that primary tumors at different subsites have different patterns of lymphatic spread, corresponding to different model parameters $\boldsymbol{\theta}$. Training a separate model for every possible subsite (ICD-10 code) would require a sufficiently large dataset for every tumor site. However, anatomically nearby locations are expected to show very similar patterns of LNL involvement. Therefore, we consider a mixture model.

Let us assume that we have a dataset $\mathbf{D}$ that is specified via the number of patients $N_{is}$ that were diagnosed in LNL involvement state $i$ and had a primary tutor in subsite $s$. Let us further assume that we want to describe this dataset using a mixture of $M$ HMMs, each with a different set of model parameters $\boldsymbol{\theta_m}$. As the generative model of the data, we assume that a patient with subsite $s$ is generated with probability $\pi_{sm}$ from model $m$. The likelihood of the dataset can then be written as

$$
P \left( \mathbf{D} \mid \boldsymbol{\theta}, \boldsymbol{\pi}\right) = \prod_s \prod_i \left[ \sum_{m=1}^M \pi_{sm} P_m \left( \mathbf{X}=\boldsymbol{\xi}_i \mid \boldsymbol{\theta}_m \right) \right]^{N_{is}}
$$ {#eq-mixture-distribution}

We now have two types of parameters, the probabilities of tutor spread for the different models $\boldsymbol{\theta_m}$, and the mixing coefficients $\pi_{sm}$. Assuming a uniform prior in the interval $[0,1]$ for all parameters, the posterior distribution over the parameters $P \left( \boldsymbol{\theta}, \boldsymbol{\pi} \mid \mathbf{D} \right)$ is given by the likelihood @eq-mixture-distribution except for a normalisation constant. In this work, we use Markov chain Monte Carlo sampling (MCMC) via the `emcee` Python package [@foreman-mackey_emcee_2013] to sample model parameters from the posterior distribution. However, $P \left( \boldsymbol{\theta}, \boldsymbol{\pi} \mid \mathbf{D} \right)$ itself is a multi-model distribution because one can permute the different models. To address this problem, we revert to an _expectation-maximization (EM)_ algorithm where we iteratively sample model parameters $\boldsymbol{\theta_m}$ using MCMC and then determine the most likely mixing coefficients.

...

The just introduced model is capable of learning one set of or distribution over parameters $\boldsymbol{\theta}$ from a cohort of patients with a primary tumor in a given subsite. If we tried to train it with a cohort consisting of patients with tumors in two very different subsites, the model would likely learn parameters that represent a compromise between the two subcohort's true parameters. This compromise might describe neither of the subcohorts' lymphatic spread patterns sufficiently well.

In such cases, mixture models are often considered. They assume the data to come from a finite mixture distribution, which -- in our particular case -- can be written as follows:

$$
P \left( \mathbf{D} \mid \boldsymbol{\Psi} \right) = \sum_{j=0}^g c_j P \left( \mathbf{D} \mid \boldsymbol{\theta}_j \right)
$$ {#eq-mixture-distribution}

Here, the $\mathbf{c} \in [0,1]^g$ is the vector of mixing proportions with $\sum_{j=0}^g c_j = 1$, while the $\boldsymbol{\Psi} = \left( \boldsymbol{\theta}_1, \ldots, \boldsymbol{\theta}_g\right)$ is the vector of all $g$ models' parameters. Note that we will implement our model such that some of the parameters in each $\boldsymbol{\theta}_j$ are shared across the $g$ components -- namely the $t_{r \rightarrow v}$ corresponding to the blue arcs in @fig-schematic-and-graph.

Let now $\mathfrak{D} = \left( \mathbf{D}_1, \ldots, \mathbf{D}_s \right)$ be a dataset consisting of $s$ subcohorts of patients. Within a subcohort $i$ we find $N_i$ patients with tumors in the same subsite. We can then introduce a _latent variable_ $\mathbf{Z}$ with a one-hot-encoding: Basically, it can take on values $\mathbf{z}_i \in \{ 0,1 \}^g$ with $z_{ij}=1$ if subcohort $i$ belongs to component $j$ and $z_{ij}=0$ else.

The latent variables are helpful in resolving the invariance of the likelihood w.r.t. permutations of the component labels, which may introduce problems, e.g. for common MCMC sampling methods. The $\mathbf{Z}$ allows us to derive two sets of interdependent equations that we may solve in an iterative fashion (see e.g. @bishop_pattern_2006 for a detailed derivation) that is commonly referred to as _expectation-maximization (EM)_ algorithm:

The first set are the probabilities of subcohort $i$ to belong to component $j$, given a set of parameters $\boldsymbol{\Psi}^\star$. These are often called the _responsibilities_:
$$
\gamma (z_{ij}) = P \left( z_{ij}=1 \mid \mathbf{D}_i , \boldsymbol{\Psi}^\star, \mathbf{c} \right) = \frac{ c_j P \left( \mathbf{D}_i \mid \boldsymbol{\theta}_j^\star \right) }{\sum_{k=0}^g c_k P \left( \mathbf{D}_i \mid \boldsymbol{\theta}_k^\star \right)}
$$

From this, we can compute new mixing proportions $c_j^\star=\sum_{i=1}^s \gamma (z_{ij}) / s$ and then infer new parameters $\boldsymbol{\Psi}^\star$ -- e.g. via MCMC sampling -- from the resulting likelihood, which is the second set:

$$
P \left( \mathbf{D} \mid \boldsymbol{\Psi}, \mathbf{c}^\star \right) = \sum_{j=0}^g c_j^\star P \left( \mathbf{D} \mid \boldsymbol{\theta}_j \right)
$$

## Implementation {#sec-implementation}

We did this in this and that fashion...

## Multi-Centric Data

For the analyses in this work, we used five datasets from four different institutions:

1. 287 oropharyngeal patients from the University of Zurich in Swizerland
2. 263 oropharyngeal patients from the Centre Léon Bérard in Fance
3. 289 oropharyngeal and oral cavity patients from the Inselspital Bern in Swizerland
4. 239 oropharyngeal and oral cavity patients from the Centre Léon Bérard in Fance
5. 162 oropharyngeal patients from the Hospital Vall d'Hebron in Spain

![Prevalence of LNL involvement stratified by subsite. The subsites are sorted in ascending order by their prevalence of involvement in LNL II. Oral cavity subsites are plotted in shaed of blue, oropharynx subsites in shades of orange.](../figures/prevalence_by_subsite.svg){#fig-prevalence-by-subsite fig-pos="b"}

The data comes in the form of CSV tables and are -- except for the last and most recent addition -- publicly available [@ludwig_detailed_2022;@ludwig_multi-centric_2023] and may be interactively explored in our **Ly**mphatic **Pro**gression e**X**plorer [LyProX](https://lyprox.org). Each row of these tables corresponds to one patient and details in which LNL metastatic involvement was found or not, according to different diagnostic and pathologic modalities.

In @fig-prevalence-by-subsite, we have plotted the prevalence of involvement in the four LNLs I, II, III, and IV stratified by the primary tumor's subsite. We only included patients with tumors in the gum, floor of mouth, other/unspecified parts of the mouth, palate, oropharynx, tosil, base of tongue, and other/unspecified parts of tongue, resulting in {{< var num_patients.total >}} patients.

# Results {#sec-results}

Works super well, of course!

![Assignment of each subsite to each of the two clusters. The fruther left a subsite, the more it is assigned to cluster A, the further right, the more to cluster B. The size of the marker (area) corresponds to the number of patients in the subsite.](../figures/cluster_assignments.svg){#fig-cluster-assignments}

# Discussion {#sec-discussion}

All that's left is for other people to bring this into clinical practice.